---
title: "Practical Machine Learning - Barbell Lifts"
author: "Óscar Martínez"
date: "October 2015"
output: html_document
---
```{r setEnvironment, echo=FALSE, warning=FALSE, message=FALSE}

# Load libraies

library(caret)
library(klaR)
library(rpart)
library(randomForest)
```
# Sipnosis
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. I'll try to quantify how well they do the activity. I will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available here: <http://groupware.les.inf.puc-rio.br/har> .

```{r load_data, echo=FALSE, warning=FALSE, message=FALSE}

# Read and clean the data
if (!file.exists("pml-training.csv")) {
download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
              destfile="pml-training.csv", method="curl")
}
trainData <- read.csv("pml-training.csv", header = TRUE, sep = ",", dec = ".", fill = TRUE,
                      na.strings=c("NA","#DIV/0!",""))
```

# Building the Model
As the goal is to predict the manner in which the subject do the activity, I'll try to classify the data of the test in the 5 differents classes explained in the paper. I test with several methods trying to get a good algorithm.
First of it all I load and clean the data provided, excluding from the analisys the 
irrelevant variables.

```{r clean_data, echo=TRUE, warning=FALSE, message=FALSE}

# delete the columns most the data is empty, and irrelevant data for the experiment
trainData   <-trainData[,-c(1:7)]
trainData <- trainData[, colSums(is.na(trainData)) < 0.8*nrow(trainData)]

```

Then, I slice the training data in train and test, in order to evaluate the models, getting the 70% of the train data for training, and the 30% for testing the models.


```{r slice_data, echo=FALSE, warning=FALSE, message=FALSE}
# Fix the seed for reproducible reseach
set.seed(123)

# Slice the data into training and testing
inTrain <- createDataPartition(y=trainData$classe, p=0.7, list=FALSE)
training <- trainData[inTrain,]
testing <- trainData[-inTrain,]

```

I build two models:

Using Rpart, getting a model with 73% of accuracy.

```{r build_model_RPART, echo=TRUE, warning=FALSE, message=FALSE}

modelRPart <- rpart(classe ~ ., data=training, method="class")

```

```{r expected_errorRPart, echo=TRUE, warning=FALSE, message=FALSE}

predictRPart <- predict(modelRPart, testing, type = "class")
confusionMatrix(predictRPart, testing$classe)

```

The result with Random Forest is pretty better. In this case, I get 99% of accuracy as we can see.

```{r build_model_RF, echo=TRUE, warning=FALSE, message=FALSE}

modelRF <- randomForest(classe ~. , data=training, method="class")

```

# Expected Out of Sample Error

Validation of Random Rorest

```{r expected_errorRF, echo=TRUE, warning=FALSE, message=FALSE}
predictRF <- predict(modelRF, testing)
confusionMatrix(testing$classe, predictRF)
```



# Submission
```{r submission, echo=FALSE, warning=FALSE, message=FALSE}
if (!file.exists("pml-testing.csv")) {
download.file(url="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
              destfile="pml-testing.csv", method="curl")
}
testData <- read.csv("pml-testing.csv", header = TRUE, sep = ",", dec = ".", fill = TRUE, na.strings=c("NA","#DIV/0!",""))


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

answers <- predict(modelRF, newdata=testData)
pml_write_files(answers)

```
